---
title: "fuel practice"
author: "Sky Biblin"
date: "2022-08-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(readxl)
library(janitor)
library(dplyr) # data summary

library(ggplot2) #plotting
library(ggthemes) #plotting
library(cowplot) #pretttier plots

library(lme4) ## run basic glmm & their fit tests
library(lmerTest) ## get p-values for lme4 & glmmADMB
library(MuMIn)
library(effects) #extracting model estimtes for plotting
library(effects) #extract model coef for plotting
# library(glmmADMB) ## zero-inflated residuals

#at http://glmm.wikidot.com/faq
#at R protool_quasi AIC ben bolker.pdf
overdisp_fun <- function(model) {
     ## number of variance parameters in 
     ##   an n-by-n variance-covariance matrix
     vpars <- function(m) {
          nrow(m)*(nrow(m)+1)/2
     }
     model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
     rdf <- nrow(model.frame(model))-model.df
     rp <- residuals(model,type="pearson")
     Pearson.chisq <- sum(rp^2)
     prat <- Pearson.chisq/rdf
     pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
     c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
     
     
}

## this function plots the residual from a GLMM for one random effect variable

## "model_output" is the name of your glmm model
## "dats" is the name of the data you used in the model
## "random_effect" is the random effect you are examining
## "graph_title" is a "text string" that is the plot's title

## EX: TestModelFit_residuals(mod.shrub.6trt, shrub, shrub$Transect, "Shrub & Transect")

TestModelFit_residuals <- function(model_output, dats, random_effect, graph_title) {

     layout(matrix(1:2, nrow=2))

     boxplot(residuals(model_output) ~ random_effect, data = dats, main = graph_title, ylab = "Residuals")

     plot(fitted(model_output), residuals(model_output), xlab = "Fitted Values", ylab = "Residuals")

     abline(h=0, lty=2)
     
     lines(smooth.spline(fitted(model_output), residuals(model_output)))

}


```
```{r}

Main_Datasheet <- read_excel("C:/Users/18052/Desktop/SJSU/Real Deal/BB_Rx--main/Main_Datasheet.xlsx")
data_clean = clean_names(Main_Datasheet)
```
fix binary data issue
```{r}
data_clean$rx <- as.factor(data_clean$rx)
unique(data_clean$rx)
```
Graph raw data
- colinearity of explanatory variables
```{r}
plot(x=data_clean$slope, y=data_clean$aspect)
plot(x=data_clean$slope, y=data_clean$canopy_cover)
plot(x=data_clean$slope, y=data_clean$heat_load)
plot(x=data_clean$slope, y=data_clean$basal_area)

plot(x=data_clean$aspect, y=data_clean$canopy_cover)
plot(x=data_clean$aspect, y=data_clean$heat_load)
plot(x=data_clean$aspect, y=data_clean$basal_area)

plot(x=data_clean$canopy_cover, y=data_clean$aspect)
plot(x=data_clean$canopy_cover, y=data_clean$heat_load)
plot(x=data_clean$canopy_cover, y=data_clean$basal_area)

plot(x=data_clean$basal_area, y=data_clean$aspect)
plot(x=data_clean$basal_area, y=data_clean$canopy_cover)
plot(x=data_clean$basal_area, y=data_clean$heat_load)

```
Model
GLMER FAQ: https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html
highlight and ctrl-shift-C to add or remove hashes
```{r}
colnames(data_clean)
```
#cwf
```{r}
data_clean$cwf <- as.integer(data_clean$cwf)

```

```{r}
mod <- glmer(data = data_clean, cwf ~ heat_load + rx + (1|site) + (1|plot_number), family = "poisson")
plot(mod)
```
```{r}
 summary(mod)
```
```{r}
r.squaredGLMM(mod)
```

Get the effect size of Rx
```{r}
ef <- effect("rx" , mod)
summary(ef)
```
Prepare to plot model predications
See https://www.r-bloggers.com/2014/08/plotting-mixed-effects-model-results-with-effects-package/
-convert effect estimates to data frame for plotting
```{r}
x <- as.data.frame(ef)
x
```
```{r}
mod.plot <- 
  ggplot(data=x, aes(x=rx, y=fit)) + 
  geom_bar(colour="black", stat="identity") +
  geom_errorbar(aes(ymin=fit-se, ymax=fit+se), width=0.4, position=position_dodge(width=0.9)) +
  xlab("") + 
  ylab("Course Woody Fuels") + 
  ggtitle("") +
  theme_tufte() +
        theme(text=element_text(size=21, family = "sans")) +
       scale_y_continuous(breaks=c(0, .1, .3, .5)) +
  scale_x_discrete(breaks=c(0, 1), 
                      labels=c("Absent", "Present")) +
    expand_limits(y=0) +
  theme(axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)))

mod.plot 

pdf('mod.plot.pdf', width = 16, height = 10)
mod.plot 
dev.off()
```
#fwf
```{r}
data_clean$fwf <- as.integer(data_clean$fwf)

```
```{r}
mod <- glmer(data = data_clean, fwf ~ heat_load + rx + (1|site) + (1|plot_number), family = "poisson")
plot(mod)
```


```{r}
 summary(mod)
```
```{r}
r.squaredGLMM(mod)
```
```{r}
ef <- effect("rx" , mod)
summary(ef)
```
```{r}
x <- as.data.frame(ef)
x
```
```{r}
mod.plot <- 
  ggplot(data=x, aes(x=rx, y=fit)) + 
  geom_bar(colour="black", stat="identity") +
  geom_errorbar(aes(ymin=fit-se, ymax=fit+se), width=0.4, position=position_dodge(width=0.9)) +
  xlab("") + 
  ylab("Fine Woody Fuels") + 
  ggtitle("") +
  theme_tufte() +
        theme(text=element_text(size=21, family = "sans")) +
       scale_y_continuous(breaks=c(0, 5, 10, 15)) +
  scale_x_discrete(breaks=c(0, 1), 
                      labels=c("Absent", "Present")) +
    expand_limits(y=0) +
  theme(axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)))

mod.plot 

pdf('mod.plot.pdf', width = 16, height = 10)
mod.plot 
dev.off()
```
#fuel depth
```{r}
data_clean$fuel_depth <- as.integer(data_clean$fuel_depth)

```

```{r}
mod <- glmer(data = data_clean, fuel_depth ~ heat_load + rx + (1|site) + (1|plot_number), family = "poisson")
plot(mod)
```
```{r}
 summary(mod)
```
```{r}
r.squaredGLMM(mod)
```
Get the effect size of Rx
```{r}
ef <- effect("rx" , mod)
summary(ef)
```
Prepare to plot model predications
See https://www.r-bloggers.com/2014/08/plotting-mixed-effects-model-results-with-effects-package/
-convert effect estimates to data frame for plotting
```{r}
x <- as.data.frame(ef)
x
```
```{r}
mod.plot <- 
  ggplot(data=x, aes(x=rx, y=fit)) + 
  geom_bar(colour="black", stat="identity") +
  geom_errorbar(aes(ymin=fit-se, ymax=fit+se), width=0.4, position=position_dodge(width=0.9)) +
  xlab("") + 
  ylab("fuel depth") + 
  ggtitle("") +
  theme_tufte() +
        theme(text=element_text(size=21, family = "sans")) +
       scale_y_continuous(breaks=c(0, .5, 1)) +
  scale_x_discrete(breaks=c(0, 1), 
                      labels=c("Absent", "Present")) +
    expand_limits(y=0) +
  theme(axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)))

mod.plot 

pdf('mod.plot.pdf', width = 16, height = 10)
mod.plot 
dev.off()
```
#Litter Depth
```{r}
data_clean$litter_depth <- as.integer(data_clean$litter_depth)

```

```{r}
mod <- glmer(data = data_clean, litter_depth ~ heat_load + rx + (1|site) + (1|plot_number), family = "poisson")
plot(mod)
```


```{r}
 summary(mod)
```
```{r}
r.squaredGLMM(mod)
```
Prepare to plot model predications
See https://www.r-bloggers.com/2014/08/plotting-mixed-effects-model-results-with-effects-package/
-convert effect estimates to data frame for plotting
```{r}
x <- as.data.frame(ef)
x
```
```{r}
mod.plot <- 
  ggplot(data=x, aes(x=rx, y=fit)) + 
  geom_bar(colour="black", stat="identity") +
  geom_errorbar(aes(ymin=fit-se, ymax=fit+se), width=0.4, position=position_dodge(width=0.9)) +
  xlab("") + 
  ylab("Litter depth") + 
  ggtitle("") +
  theme_tufte() +
        theme(text=element_text(size=21, family = "sans")) +
       scale_y_continuous(breaks=c(0, .5, 1)) +
  scale_x_discrete(breaks=c(0, 1), 
                      labels=c("Absent", "Present")) +
    expand_limits(y=0) +
  theme(axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)))

mod.plot 

pdf('mod.plot.pdf', width = 16, height = 10)
mod.plot 
dev.off()
```
#aerial fuels
 
```{r}

Main_Datasheet <- read_excel("C:/Users/18052/Desktop/SJSU/Real Deal/BB_Rx--main/Main_Datasheet.xlsx")
data_clean = clean_names(Main_Datasheet)

```
fix binary data issue
```{r}
data_clean$rx <- as.factor(data_clean$rx)
unique(data_clean$rx)
```

```{r}
data_clean$sum_top_kill <- as.integer(data_clean$sum_top_kill)
```
```{r}
mod <- glmer(data = data_clean, sum_top_kill ~ heat_load + rx + (1|site) + (1|plot_number), family = "poisson")
plot(mod)
```
```{r}
 summary(mod)
```
```{r}
r.squaredGLMM(mod)
```
Get the effect size of Rx
```{r}
ef <- effect("rx" , mod)
summary(ef)
```
Prepare to plot model predications
See https://www.r-bloggers.com/2014/08/plotting-mixed-effects-model-results-with-effects-package/
-convert effect estimates to data frame for plotting
```{r}
x <- as.data.frame(ef)
x
```
```{r}
mod.plot <- 
  ggplot(data=x, aes(x=rx, y=fit)) + 
  geom_bar(colour="black", stat="identity") +
  geom_errorbar(aes(ymin=fit-se, ymax=fit+se), width=0.4, position=position_dodge(width=0.9)) +
  xlab("") + 
  ylab("Aerial Fuels") + 
  ggtitle("") +
  theme_tufte() +
        theme(text=element_text(size=21, family = "sans")) +
       scale_y_continuous(breaks=c(0, 5, 10, 15)) +
  scale_x_discrete(breaks=c(0, 1), 
                      labels=c("Absent", "Present")) +
    expand_limits(y=0) +
  theme(axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)))

mod.plot 

pdf('mod.plot.pdf', width = 16, height = 10)
mod.plot 
dev.off()
```
#canopy retained 

```{r}
data_clean$percent_canopy_retained <- as.integer(data_clean$percent_canopy_retained)
```
```{r}
mod <- glmer(data = data_clean, percent_canopy_retained ~ heat_load + rx + (1|site) + (1|plot_number), family = "poisson")
plot(mod)
```
```{r}
 summary(mod)
```
```{r}
r.squaredGLMM(mod)
```
```{r}
ef <- effect("rx" , mod)
summary(ef)
```